Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2017-08-15 paper.tex
Changes from Kang-Kuen

\begin{figure*}
  \includegraphics[height=4.5cm]{fig1_combined.pdf}
  \caption{Single Na atom Raman sideband cooling scheme and sequence. (A)
    The Raman transitions between $|2,-2;n\rangle$ and $|1,-1;n+\Delta n\rangle$ have a one-photon detuning $\Delta=25$ GHz below the $3^2S_{1/2}$ to $3^2P_{3/2}$ transition. Two-photon detuning, $\delta$, is defined relative to the $\Delta n=0$ carrier transition. For optical pumping, we use two $\sigma^-$ polarized transitions, one to pump the atom state out of $|1,-1\rangle$ via $3^2P_{3/2}$ and one to pump atoms out of $|2,-1\rangle$ via $3^2P_{1/2}$
     to minimize heating of the $|2,-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.  Raman beams R1 and R4 address the radial $x$-mode. R1 and R2 address the radial $y$-mode.  R3 and R4 address the axial $z$-mode, where the beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.  %, which also couples to the radial $y$ mode.    The radial $x$-mode is addressed by turning on beams R1 and R4, $y$ is addressed by R1 and R2, and $z$ is addressed by R3 and R4.  In this geometry, the radial mode directions coincide with the respective $k$-vector differences.  The ``axial'' Raman beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.
    (C) Schematic of the cooling pulse sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ sideband pulses.
    The four axial pulses address two sideband orders.
    The two pulses in each radial direction either address $\Delta n=-2$ and $\Delta n=-1$
    or have different durations to drive $\Delta n=-1$, at the end of the cooling sequence when most of the population is below $n=3$.
    The Raman cooling and spectroscopy pulses have Blackman envelopes~\cite{Kasevich1992}
    to reduce off-resonant coupling,
    while the measurement Rabi pulses in Fig.~\ref{f-radial}(B,C) and \ref{f-axial}(B,C)
    have square envelopes to simplify analysis.
    \label{f-setup}}
\end{figure*}

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (total duration $100$ ms) are applied
along three axes with cooling beginning on the radial second order and axial eighth order.
To characterize the single atom thermal state before and after cooling,
we perform Raman sideband thermometry~\cite{Monroe1995, Meekhof1996}.
For the more tightly confined (non-degenerate) radial directions,
we observe clear $\Delta n=1$, $\Delta n=-1$, and $\Delta n=-2$ sidebands before RSC, as is shown in Fig.~\ref{f-radial}A.
After RSC, the $\Delta n=-1$ and $\Delta n=-2$ sidebands on both radial axes are strongly reduced.
%It is important to note that the simple sideband thermometry formula, which predicts the ratio of the cooling and heating sideband heights to be $\bar n / (\bar n + 1)$~\cite{Monroe1995}, assumes that the coupling strength for different motional states is proportional to $\sqrt{n}$. However, outside of the LD regime, the coupling strength rapidly deviates from this simple scaling rule, as shown in Figure \ref{f-ld}A. Since the Raman spectroscopy is performed with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to measure $\bar n$.
The formula that equates the ratio of $\Delta n=-1$ and $\Delta n=1$ sideband heights to $\bar{n}/(\bar{n}+1)$ ($\bar{n}$ is the mean motional quantum number of the assumed thermal state) was derived in the LD regime~\cite{Monroe1995}. However, it is also valid outside the LD regime.  The resulting $\bar{n}_x=0.09(3)$ and $\bar{n}_y=0.05(2)$ correspond to ground-state fractions of 92(2)\% and 95(2)\%,  in agreement with fitted values of 89(2)\% and 92(2)\% from the Rabi flopping curves~ \cite{Meekhof1996} in Fig.~\ref{f-radial}B and \ref{f-radial}C.
%Instead, given the absence of the second order cooling sideband, we assume that the atoms are cold enough to have stronger coupling to the first order cooling sideband than the second order, i.e. below where the matrix element for the two orders cross, in which case the lower bound of the ground state population can be computed from the heights of the sidebands by assuming all of the excitations are in the $n=1$ state. In doing this, we found a ground state population of $90(2)$\% and $94(3)$\% along the two radial directions. We also use measurements of Rabi flopping on the carrier (Fig. \ref{f-radial}B) and first-order heating(Fig. \ref{f-radial}C) sidebands to verify our assumption about the distribution and the ground state population~\cite{Meekhof1996}. The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling which yields $89(2)$\% and $92(2)$\% and show good agreement with the population extracted from the sideband thermometry.
The initial temperature of $70\,\mu$K before RSC is obtained
from similar fits. %the motional-state decohered Rabi flopping.

>>>>

\begin{figure*}
  \includegraphics[height=4.5cm]{fig1_combined.pdf}
  \caption{Single Na atom Raman sideband cooling scheme and sequence. (A)
    The Raman transitions between $|2,-2;n\rangle$ and $|1,-1;n+\Delta n\rangle$ have a one-photon detuning $\Delta=25$ GHz below the $3^2S_{1/2}$ to $3^2P_{3/2}$ transition. Two-photon detuning, $\delta$, is defined relative to the $\Delta n=0$ carrier transition. For optical pumping, we use two $\sigma^-$ polarized transitions, one to pump the atom state out of $|1,-1\rangle$ via $3^2P_{3/2}$ and one to pump atoms out of $|2,-1\rangle$ via $3^2P_{1/2}$
     to minimize heating of the $|2,-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.  Raman beams R1 and R4 address the radial $x$-mode. R1 and R2 address the radial $y$-mode.  R3 and R4 address the axial $z$-mode, where the beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.  %, which also couples to the radial $y$ mode.    The radial $x$-mode is addressed by turning on beams R1 and R4, $y$ is addressed by R1 and R2, and $z$ is addressed by R3 and R4.  In this geometry, the radial mode directions coincide with the respective $k$-vector differences.  The ``axial'' Raman beams also couple to radial motion, but this coupling can be neglected when the atoms is cooled to the ground state of motion.
    (C) Schematic of the cooling pulse sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ sideband pulses.
    The four axial pulses address two sideband orders.
    The two pulses in each radial direction either address $\Delta n=-2$ and $\Delta n=-1$
    or have different durations to drive $\Delta n=-1$, at the end of the cooling sequence when most of the population is below $n=3$.
    The Raman cooling and spectroscopy pulses (excluding $\Delta n=-8$ to $-5$ pulses) have Blackman envelopes~\cite{Kasevich1992}
    to reduce off-resonant coupling,
    while the measurement Rabi pulses in Fig.~\ref{f-radial}(B,C) and \ref{f-axial}(B,C)
    have square envelopes to simplify analysis.
    \label{f-setup}}
\end{figure*}

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (total duration $100$ ms) are applied
along three axes with cooling beginning on the radial second order and axial eighth order.
The full RSC pulse sequence is given in~\footnote{See supplemental material}.
To characterize the single atom thermal state before and after cooling,
we perform Raman sideband thermometry~\cite{Monroe1995, Meekhof1996}.
For the more tightly confined (non-degenerate) radial directions,
we observe clear $\Delta n=1$, $\Delta n=-1$, and $\Delta n=-2$ sidebands before RSC, as is shown in Fig.~\ref{f-radial}A.
After RSC, the $\Delta n=-1$ and $\Delta n=-2$ sidebands on both radial axes are strongly reduced.
%It is important to note that the simple sideband thermometry formula, which predicts the ratio of the cooling and heating sideband heights to be $\bar n / (\bar n + 1)$~\cite{Monroe1995}, assumes that the coupling strength for different motional states is proportional to $\sqrt{n}$. However, outside of the LD regime, the coupling strength rapidly deviates from this simple scaling rule, as shown in Figure \ref{f-ld}A. Since the Raman spectroscopy is performed with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to measure $\bar n$.
The formula that equates the ratio of $\Delta n=-1$ and $\Delta n=1$ sideband heights to $\bar{n}/(\bar{n}+1)$ ($\bar{n}$ is the mean motional quantum number of the assumed thermal state) was derived in the LD regime~\cite{Monroe1995}. However, it is also valid outside the LD regime.  The resulting $\bar{n}_x=0.09(3)$ and $\bar{n}_y=0.05(2)$ correspond to ground-state fractions of 92(2)\% and 95(2)\%,  in agreement with fitted values of 89(2)\% and 92(2)\% from the Rabi flopping curves~ \cite{Meekhof1996} in Fig.~\ref{f-radial}B and \ref{f-radial}C.
%Instead, given the absence of the second order cooling sideband, we assume that the atoms are cold enough to have stronger coupling to the first order cooling sideband than the second order, i.e. below where the matrix element for the two orders cross, in which case the lower bound of the ground state population can be computed from the heights of the sidebands by assuming all of the excitations are in the $n=1$ state. In doing this, we found a ground state population of $90(2)$\% and $94(3)$\% along the two radial directions. We also use measurements of Rabi flopping on the carrier (Fig. \ref{f-radial}B) and first-order heating(Fig. \ref{f-radial}C) sidebands to verify our assumption about the distribution and the ground state population~\cite{Meekhof1996}. The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling which yields $89(2)$\% and $92(2)$\% and show good agreement with the population extracted from the sideband thermometry.
The initial temperature of $70\,\mu$K before RSC is obtained
from similar fits. %the motional-state decohered Rabi flopping.
